Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-10 paper.qmd
Updated Figure 9 (a)

@@

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)

  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

@@

# Define bin width and starting point
bin_width <- hb_obj_two_curvy2$a1
start_point <- bin_width/2

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  #geom_jitter() +
  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
  #                    labels = c(
  #                      expression(a[1]),
  #                      expression(2*a[1]),
  #                      expression(3*a[1]),
  #                      expression(4*a[1]),
  #                      expression(5*a[1]),
  #                      expression(6*a[1]),
  #                      expression(7*a[1]),
  #                      expression(8*a[1])
  #                    )) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
           linetype="solid",
           color = "#bdbdbd",
           linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
  count(bin) |>
  mutate(prop = n/sum(n)) |>
  head(4)

# distance_two_curvy_dist <- distance_two_curvy_dist |>
#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small

text_df <- tibble(
  x = c(1.9, 3.4),
  y = c(0.5, 0.5),
  text = c("e1", "d1")
)

distance_hist <- ggplot(
  distance_two_curvy_dist_summary,
  aes(x = as.numeric(bin),
      y = prop)) +
  geom_line() +
  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1])
                     )) +
  geom_vline(xintercept = 2,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 3.5,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_text(
    data = text_df,
    aes(
      x = x,
      y = y,
      label = text
    ),
    size = 9
  ) +
  ylab("Proportion of the number of edges") +
  xlab(expression(d^{(2)})) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

# distance_hist <- ggplot(
#   distance_two_curvy_dist, aes(x = distance)) +
#     geom_histogram(
#       aes(y = after_stat(count) / sum(after_stat(count))), 
#       binwidth = bin_width) +
#   scale_x_continuous(breaks = bin_labels, 
#                      labels = c(
#                        expression(a[1]),
#                        expression(2*a[1]),
#                        expression(3*a[1]),
#                        expression(4*a[1]),
#                        expression(5*a[1]),
#                        expression(6*a[1]),
#                        expression(7*a[1]),
#                        expression(8*a[1])
#                      )) +
#   geom_vline(xintercept = start_point + 2*bin_width,
#              linetype="solid",
#              color = "#bdbdbd",
#              linewidth=1) +
#   ylab("Proportion of the number of edges") +
#   xlab(expression(d^{(2)})) +
#   #ggtitle("(d)") +
#   theme_minimal() +
#   theme(aspect.ratio = 0.75,
#         panel.border = element_rect(fill = 'transparent'),
#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
#         panel.grid.major.x = element_blank(),
#         axis.ticks.x = element_line(),
#         axis.ticks.y = element_line())

>>>>

```{r}
#| label: dist-plot

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)

  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

# Define bin width and starting point
bin_width <- hb_obj_two_curvy$a1
start_point <- bin_width/2

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

@@
